---
title: "HW 6"
author: "Yoonseo Mok"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DBI)     
library(parallel)
library(future)
library(data.table)
```

```{r}
# Import the SQLite database of the Lahman data
lahman <- dbConnect(RSQLite::SQLite(), "lahman_1871-2022.sqlite")

# Reading in fielding data
Fielding=dbGetQuery(lahman, "SELECT * FROM Fielding")
# Calculating Range Factor
Fielding$RF=3*(Fielding$PO+Fielding$A)/Fielding$InnOuts
# Data cleaning
fielding=na.omit(Fielding[, c("teamID", "RF")])
fielding=fielding[!is.infinite(fielding$RF),]

# Calculating the average RF for each team
point_est=aggregate(fielding$RF,by=list(fielding$teamID), FUN =mean)
colnames(point_est)=c("teamID", "meanRF")
```

# A
```{r}
fielding_dat=data.table(fielding)

# Bootstrap using data.table
f = function(data) {
  tmp=data[,.SD[sample(.N, replace=TRUE)], by=teamID]
  return(tmp[,.(meanRF = mean(RF)), by=teamID])
}

# Checking the function
f(fielding_dat)

# Measuring time
system.time(f(fielding_dat))
```

# A1: Without any parallel processing
```{r}
reps=1000

# Using lapply
system.time({
  result1=lapply(seq_len(reps), function(x) f(fielding_dat))
})
```

# A2: Using parallel processing with the parallel package
```{r}
# Using mcapply 
system.time({
  result2=mclapply(seq_len(reps), function(x) f(fielding_dat), mc.cores=5)
})
```

# A3: Using futures with the future package
```{r}
# Using future
plan(multisession)
system.time({
  result3=lapply(seq_len(reps), function(x) {
    future(f(fielding_dat),seed=TRUE)
  })
  result3 =lapply(result3,value)
})
```

# B 
```{r}
# Generating a table showing the estimated RF and associated SE from the three approaches.
tablesum <- function(results) {
  bind=rbindlist(results)
  summary=bind[, .(meanRF=mean(meanRF), SE=sd(meanRF)), by=teamID]
  return(summary)
}

# Showing the results
print(tablesum(result1), n = 1e3)
print(tablesum(result2), n = 1e3)
print(tablesum(result3), n = 1e3)
```

# C
Without Parallel Processing (lapply) it took 11.7 seconds for 1000 simulations
Using parallel (mclapply) it took 2.6 seconds for 1000 simulations
Using future Package, it took 72.8 seconds for 1000 simulations

Using parallel (mclapply) is the best choice for this problem. It is the fastest.

But all three methods resulted similar mean estimates and standard errors