---
title: "HW1"
output: pdf_document
author: Yoonseo Mok
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Problem 1

# A
```{r}
# Importing wine dataset
wine <- read.csv("~/Desktop/Fall 2024/STATS 506/HW1/wine/wine.data", header=FALSE)
names(wine)<-c("Class","Alcohol", "Malic acid", "Ash", "Alcalinity of ash", "Magnesium", "Total phenols",
               "Flavanoids", "Nonflavanoid phenols", "Proanthocyanins", 
 	             "Color intensity", "Hue", "OD280/OD315 of diluted wines", "Proline")
```

# B
```{r}
# Checking how many wines within each class
table(wine$Class)
```
Class 1: 59, Class 2:71, Class 3: 48\
It is the same number as reported in "wine.names"

# C-1
```{r}
# Looking at correlation
cor(wine$Alcohol,wine$`Color intensity`)
```
The correlation is 0.55

# C-2
```{r}
# Looking at correlation by each class between alcohol content and color intensity
wine %>% group_by(Class) %>% summarise(cor=cor(Alcohol,`Color intensity`))
```
Class 1 has the highest correlation and class 2 has the lowest correlation between alcohol content and color intensity

# C-3
```{r}
# Alcohol content that has the highest color intensity
wine$Alcohol[which.max(wine$`Color intensity`)]
```
The alcohol content of the wine with the highest color intensity is 14.34

# C-4
```{r}
# Comparing proanthocyanins to ash
tmp=wine[wine$Proanthocyanins>wine$Ash,]
nrow(tmp)/nrow(wine)*100
```
8.43% of wines had a higher content of proanthocyanins compare to ash

# D
```{r}
tmp2=matrix(NA,4*(ncol(wine)-1),2)
colnames(tmp2)=c("Variable","Average")
for(i in 2:ncol(wine)){
  tmp2[4*i-7,1]=paste0(names(wine)[i],"-Overall")
  tmp2[4*i-6,1]=paste0(names(wine)[i],"-Class 1")
  tmp2[4*i-5,1]=paste0(names(wine)[i],"-Class 2")
  tmp2[4*i-4,1]=paste0(names(wine)[i],"-Class 3")

  tmp2[4*i-7,2]=round(mean(wine[,i],na.rm=T),2)
  tmp2[4*i-6,2]=round(mean(wine[wine$Class %in% 1,i],na.rm=T),2)
  tmp2[4*i-5,2]=round(mean(wine[wine$Class %in% 2,i],na.rm=T),2)
  tmp2[4*i-4,2]=round(mean(wine[wine$Class %in% 3,i],na.rm=T),2)
}

kable(tmp2) %>% kable_styling("striped", full_width = F) 
```

# E
```{r}
# Class 1 vs Class 2
class12=subset(wine,wine$Class %in% c(1,2))
t.test(class12$`Total phenols`~class12$Class)
```

```{r}
# Class 1 vs Class 3
class13=subset(wine,wine$Class %in% c(1,3))
t.test(class13$`Total phenols`~class13$Class)
```

```{r}
# Class 2 vs Class 3
class23=subset(wine,wine$Class %in% c(2,3))
t.test(class23$`Total phenols`~class23$Class)
```
The p-values of all three t-test are significant and show that the level of phenols differs across the three classes. The mean of phenols in group 1 vs group 2 vs group 3 are significantly different.

# Problem 2

# A
```{r}
# Importing dataset
AskAManager <- read.csv("~/Desktop/Fall 2024/STATS 506/HW1/AskAManager.csv")
```

# B
```{r}
# Cleaning column name
names(AskAManager)=c("ID","Timestamp","Age","Industry","Jobtitle","Jobtitle_add","Salary","Salary_add","Currency","Currency_Other","Context","Country","State","City","Num years_overall","Num years_field","Education","Gender","Race")
```

# C
```{r}
# Only USD
USDonly=AskAManager[AskAManager$Currency %in% "USD",]
nrow(AskAManager)
nrow(USDonly)
```
Comparing the number of observations, before it was 20862 and after restricting we now have 23374

# D
```{r}
# No one starts working before age 18
no18=USDonly[!(USDonly$Age %in% "under 18"),]

# Taking the midpoint of age
no18$Age_rev=NA
no18$Age_rev[no18$Age %in% "18-24"]=21
no18$Age_rev[no18$Age %in% "25-34"]=30
no18$Age_rev[no18$Age %in% "35-44"]=40
no18$Age_rev[no18$Age %in% "45-54"]=50
no18$Age_rev[no18$Age %in% "55-64"]=60
no18$Age_rev[no18$Age %in% "65 or over"]=65

# Taking the midpoint of overall experience of working 
no18$`Num years_overall_rev`=NA
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "1 year or less"]=0.5
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "2 - 4 years"]=3
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "5-7 years"]=6
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "8 - 10 years"]=9
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "11 - 20 years"]=15
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "21 - 30 years"]=25
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "31 - 40 years"]=35
no18$`Num years_overall_rev`[no18$`Num years_overall` %in% "41 years or more"]=45

# Taking the midpoint of field experience of working 
no18$`Num years_field_rev`=NA
no18$`Num years_field_rev`[no18$`Num years_field` %in% "1 year or less"]=0.5
no18$`Num years_field_rev`[no18$`Num years_field` %in% "2 - 4 years"]=3
no18$`Num years_field_rev`[no18$`Num years_field` %in% "5-7 years"]=6
no18$`Num years_field_rev`[no18$`Num years_field` %in% "8 - 10 years"]=9
no18$`Num years_field_rev`[no18$`Num years_field` %in% "11 - 20 years"]=15
no18$`Num years_field_rev`[no18$`Num years_field` %in% "21 - 30 years"]=25
no18$`Num years_field_rev`[no18$`Num years_field` %in% "31 - 40 years"]=35
no18$`Num years_field_rev`[no18$`Num years_field` %in% "41 years or more"]=45

# Now subtract age from overall experience of working
no18$startwork=NA
no18$startwork=no18$Age_rev-no18$`Num years_overall_rev`

# Exclude those who have negative start work value
no18=no18[no18$startwork>=0,]

# Now subtract age from field experience of working
no18$startwork_field=NA
no18$startwork_field=no18$Age_rev-no18$`Num years_field_rev`

# Exclude those who have negative start work in field value
no18=no18[no18$startwork_field>=0,]

nrow(no18)
```
Eliminating any rows for which their age, years of experience in their field, and years of experience total are impossible, I have 23340 rows.

# E
```{r}
# First adding annual salary and bonus
no18$finalsalary=rowSums(no18[,c("Salary", "Salary_add")], na.rm=TRUE)

# Find the outlier value based on IQR
low=mean(no18$finalsalary)-1.5*IQR(no18$finalsalary)
up=mean(no18$finalsalary)+1.5*IQR(no18$finalsalary)

# Eliminating any outlier(extreme) income
incomeokay=no18[no18$finalsalary<=up & no18$finalsalary>low,]
nrow(incomeokay)
```
I considered the extreme value of income using 1.5*IQR. Any income less than $13000 and higher than 199010 is considered as outlier based on our data. After excluding those people I have 21643 number of observation in my data

# Problem 3
# A
```{r}
#' Identify if it is a palindrome and shows the reverse 
#' @param x A number.
#' @returns A list if it is palindromic and the reverse of x
#' @examples isPalindromic(199) 

isPalindromic=function(x){
  x=as.character(x)
  
  if(x<0){
    stop("It is a negative integer")
  }
  
  if(substr(x,nchar(x),nchar(x))==0){
    stop("The integer ends with 0")
  }
  
  reversed=stringi::stri_reverse(x)
  
  if(nchar(x) %% 2 == 0){ # If x is even number
    i=nchar(x)
    
    firsthalfchar=substr(x,1,i/2)
    lasthalfchar=substr(x,(i/2)+1,i)
    
    if(stringi::stri_reverse(firsthalfchar) == lasthalfchar){
      isPalindromic=TRUE
    } else {
      isPalindromic=FALSE
    }
    
  } else { # If x is odd number
    i=nchar(x)
    
    firsthalfchar=substr(x,1,floor(i/2))
    lasthalfchar=substr(x,ceiling(i/2)+1,i)
    
    if(stringi::stri_reverse(firsthalfchar) == lasthalfchar){
      isPalindromic=TRUE
    } else {
      isPalindromic=FALSE
    }
  }
    return(as.list(data.frame(isPalindromic,reversed)))
}

isPalindromic(728827)
isPalindromic(39951)
```

# B
```{r}
#' Finds the next palindromic number strictly greater than the input
#' @param x A number.
#' @returns A numeric vector
#' @examples nextPalindrome(199) 

nextPalindrome=function(x){
  x=as.character(x)
  
  if(x<0){
    stop("It is a negative integer")
  }
  
  if(nchar(x) %% 2 == 0){ # If x is even number
    i=nchar(x)
    
    firsthalfchar=substr(x,1,i/2)
    
    if(as.numeric(paste0(firsthalfchar,stringi::stri_reverse(firsthalfchar)))>x){
      # If the first half numbers and the reverse of that is greater than x then,
      nextvalue=as.numeric(paste0(firsthalfchar,stringi::stri_reverse(firsthalfchar)))
    } else {
      # Else, take the number that is right after the half (first number of the second half)
      # and replace the last number in the first half
      # then, reverse that
      takelastvalue=substr(x,(i/2)+1,(i/2)+1)
      takelastvalueadd=paste0(substr(firsthalfchar,1,(i/2)-1),takelastvalue)
      nextvalue=as.numeric(paste0(takelastvalueadd,stringi::stri_reverse(takelastvalueadd)))
    }
    
  } else { # if x is odd number
    i=nchar(x)
    
    firsthalfchar=substr(x,1,floor(i/2))
    if(as.numeric(paste0(firsthalfchar,substr(x,floor(i/2)+1,floor(i/2)+1),stringi::stri_reverse(firsthalfchar)))>x){
      # If the first half numbers with the middle number and the reverse of that is greater than x then,
      nextvalue=as.numeric(paste0(firsthalfchar,substr(x,floor(i/2)+1,floor(i/2)+1),stringi::stri_reverse(firsthalfchar)))
    } else {
      # similar logic with even number
      # +1 with the middle number instead
      # we also need to take account when the middle number is 9 
      # because 9+1=10
      takelastvalue=substr(x,floor(i/2)+1,floor(i/2)+1)
      if(takelastvalue<9){ 
        takelastvalueadd=paste0(substr(firsthalfchar,1,floor(i/2)+1),(as.numeric(takelastvalue)+1))
        nextvalue=as.numeric(paste0(takelastvalueadd,stringi::stri_reverse(firsthalfchar)))
      } else {
        takelastvalueadd=paste0(as.numeric(substr(x,1,ceiling(i/2)))+1)
        nextvalue=as.numeric(paste0(takelastvalueadd,stringi::stri_reverse(takelastvalueadd)))
      }
    }
    
  }
    return(nextvalue)
}

nextPalindrome(391)
nextPalindrome(9928)
nextPalindrome(19272719)
nextPalindrome(109)
nextPalindrome(2)
```
