---
title: "HW2"
author: "Yoonseo Mok"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github link
https://github.com/mokys1213/STATS-506-FA-2024/blob/main/HW2/HW2.pdf

# Problem 1
# A
## Version 1
```{r}
#' Version 1: Implement this game using a loop
#'
#' @param n The number of dice to roll.
#' @return Total winnings

v1 <- function(n) {
  tot_win <- 0
  for (i in 1:n) {
    rollnumb <- sample(1:6, 1)  
    if (rollnumb %in%  3 | rollnumb %in% 5) {
      tot_win <- tot_win + 2 * rollnumb - 2
    } else {
      tot_win <- tot_win - 2
    }
  }
  return(tot_win)
}
```

## Version 2
```{r}
#' Version 2: Implement this game using built-in R vectorized functions.
#'
#' @param n The number of dice to roll.
#' @return Total winnings

v2 <- function(n) {
  rollnumb <- sample(1:6, n,replace = T)  
  win=ifelse(rollnumb %in% c(3,5),2*rollnumb-2,-2)
  total_win=sum(win)
  return(total_win)
}
```

## Version 3
```{r}
#' Version 3: Implement this by rolling all the dice into one 
#' and collapsing the die rolls into a single table()
#'
#' @param n The number of dice to roll.
#' @return Total winnings

v3 <- function(n) {
  rollnumb <- sample(1:6, n, replace = TRUE)  
  rollnumb_table <- table(factor(rollnumb, levels = 1:6))
  
  tot_win <- sum((rollnumb_table[3] * (3*2 - 2)) + (rollnumb_table[5]* (5*2 - 2)), na.rm = TRUE)
  tot_lose <- sum(rollnumb_table[c(1,2,4,6)] * (-2), na.rm = TRUE)

  return(tot_win + tot_lose)
}

```

## Version 4
```{r}
#' Version 4: Implement this game by using one of the “apply” functions.
#'
#' @param n The number of dice to roll.
#' @return Total winnings

v4 <- function(n) {
  rollnumb <- sample(1:6, n, replace = TRUE)  
  winorlose <- sapply(rollnumb, function(x) {
    if (x %in%  3 | x %in% 5) {
      return(2*x-2)
    } else {
      return(-2)
    }
   }
  )
  return(sum(winorlose))
}
```

# B
```{r}
# Demonstrate that all versions work
v1(3)
v1(3000)
v2(3)
v2(3000)
v3(3)
v3(3000)
v4(3)
v4(3000)
```
All versions work

# C
```{r}
# Demonstrate that the four versions give the same result. 
set.seed(123)
v1(3)
set.seed(123)
v1(3000)
set.seed(123)
v2(3)
set.seed(123)
v2(3000)
set.seed(123)
v3(3)
set.seed(123)
v3(3000)
set.seed(123)
v4(3)
set.seed(123)
v4(3000)
```
All four versions give the same result. 

# D
```{r, warning=FALSE}
library(microbenchmark)

# Demonstrate the speed of the implementations
# Low input of 1,000 
x1<-1000
microbenchmark(v1(x1), v2(x1),v3(x1),v4(x1))

# High input of 100000
x2<-100000
microbenchmark(v1(x2), v2(x2),v3(x2),v4(x2))
```
Comparing the performance with low input of 1000, Version 2 function where we implemented the game using built-in R vectorized functions has the fastest performance. 
Comparing the performance with high input of 100000, Version 3 function where we Implemented the dice rolls into a single table() has the fastest performance. While version 3 is the fastest, version 2 function performance is similar. 
In both comparison, Version 1 function where we implemented the game using loop performs the slowest.

# E
```{r}
# Evidence based upon a Monte Carlo simulation
simnumb <- 1000000
set.seed(123)

# Generate total winnings using version 2 function
total_net_result = v2(simnumb)

# Calculate average net result per game
average_net_result <- total_net_result / simnumb
average_net_result

# Calculating using probability (just another way of providing evidence)
1/6*4+1/6*8+4/6*(-2)
```
Based on the monte carlo simulation, we can see that we win 0.66. Therefore, we can conclude that this is a fair game since we are not loosing anything at least. 

# Problem 2
```{r}
# Importing the dataset
cars=read.csv("cars.csv")
```

# A
```{r}
# Rename the columns 
names(cars)=c("Height","Length","Width","Driveline","EngineType","Hybrid","NumbForwardGears",
              "Transmission","Citympg","FuelType","Highwaympg","Class","ID","Make","ModelYear","Year",
              "Horsepower","Torque")
```

# B
```{r}
# Restrict the data to cars whose Fuel Type is “Gasoline”.
cars=cars[cars$FuelType %in% "Gasoline",]
```

# C
```{r}
# Checking highway mileage variable distribution
summary(cars$Highwaympg)
hist(cars$Highwaympg)

# Log transformation of highway mpg variable
cars$Highwaympg_log=log(cars$Highwaympg)
hist(cars$Highwaympg_log)
```
Highway gas mileage variable is slightly skewed. Although it is not extremely skewed, a log transformation could be used to normalize the highway mpg variable.

# D
```{r}
# Fit a linear regression model
model1=lm(Highwaympg_log ~ Torque+Horsepower+Height+Length+Width+factor(Year), data = cars)
summary(model1)
```
exp(-2.294e-03)=0.9977086
For every one-unit increase in the torque variable, highway gas mileage decreases by a factor of about 0.9977086, holding horsepower, all three dimensions of the car, and the year the car was released as constant.

# E
```{r}
# Fit a linear regression model with interaction
model2=lm(Highwaympg_log ~ Torque*Horsepower+Height+Length+Width+factor(Year), data = cars)
summary(model2)

library(interactions)

# Create interaction plot
interact_plot(model2, pred = Torque, modx = Horsepower, 
              modx.values = c(170, 250, 330), at = list(Year = 2010), data=cars)
interact_plot(model2, pred = Torque, modx = Horsepower, 
              modx.values = c(170, 250, 330), at = list(Year = 2011), data=cars)
interact_plot(model2, pred = Torque, modx = Horsepower, 
              modx.values = c(170, 250, 330), at = list(Year = 2012), data=cars)
```

# F
```{r}
# Creating design matrix
X <- model.matrix( ~ Torque + Horsepower + Height + Length + Width + factor(Year), data = cars)
Y <- cars$Highwaympg_log

# Calculate the beta coefficients using design matrix
solve(t(X) %*% X) %*% t(X) %*% Y

```
It gives out the same results as lm did prior

